%!TEX root = ../lectures.tex

\section{Graded modules \& algebras}\label{lecture:graded-algebra}
In areas incorporating homological algebra, including homological algebra itself, one is commonly concerned with chain complexes of modules over a ring.
However, often these chain complexes admit a bit more structure than merely being chain complexes: they can in many cases admit a kind of multiplication.
Forgetting the differential, one has merely \emph{graded} modules.

\subsection{Graded objects}
Let us fix a commutative ring \(\Bbbk\). Here, we use \(\Bbbk\) to denote a \emph{ring} instead of a \emph{field} for historical reasons.
Commonly, a (\(\Z\)-)\emph{graded} \(\Bbbk\)\emph{-module} is said to be a \(\Bbbk\)-module \(M\) together with a decomposition
\[ M \cong \coprod_{i\in\Z}M^i \]
of \(M\) as a \(\Z\)-indexed coproduct of \(\Bbbk\)-modules---the \emph{grading} on \(M\). It is not hard to see that this datum is really equivalent to \(M\), and
in particular, it makes sense to make the following definition.
\begin{definition}
	Let \(\calC\) be a category. The category of \(\Z\)-graded objects in \(\calC\) is the functor category
	\[ \Fun(\Z,\calC) \cong \prod_{\Z}\calC, \]
	where \(\Z\) is regarded as a discrete category. In particular, we obtain the category \(\gMod_\Bbbk\) of \emph{graded} \(\Bbbk\)\emph{-modules}
	\[ \gMod_{\Bbbk} \coloneq \prod_{\Z}\Mod_{\Bbbk}. \]
	For \(x\in\Fun(\Z,\calC)\), we call \(x^i\) the \emph{degree \(i\) piece of \(x\).} For \(M\in\gMod_\Bbbk\), we say that an element
	of \(M^i\) is \emph{homogeneous of degree \(i\).}
\end{definition}
\begin{remark}
	We will essentially always be working with \(\Z\)-graded objects, so henceforth we will drop bothering to specify that.
\end{remark}
\begin{remark}
	One could have defined graded \(\Bbbk\)-modules as being pairs \((M,(M^i)_{i\in\Z})\), with morphisms being the \(\Bbbk\)-linear maps
	which preserve the grading. It is easily seen that this results in an equivalent category.
\end{remark}

\begin{terminology}
	Let \(M,N\in\gMod_\Bbbk\). A morphism \(M\to N\) in \(\gMod_\Bbbk\) is said to be a \emph{graded morphism of degree \(0\),}
	or otherwise a \emph{strict morphism.}
\end{terminology}

Let us translate some of the terminology standard to graded objects to our setting. In particular, let us detail how one encodes morphisms
of non-zero degree in this formalism.

\begin{definition}
	Let \(\calC\) be a category. The automorphism \(k\mapsto k+1\) of \(\Z\) induces the automorphism
	\[ (1)\!:\Fun(\Z,\calC)\to\Fun(\Z,\calC),\quad x = (x^i)_{i\in\Z} \mapsto (x^{i+1})_{i\in\Z} \eqcolon x(1). \]
	of categories of graded objects in \(\calC\). We denote by \((-1)\) the inverse of this functor, and thus produce functors
	\((i)\) for all \(i\in\Z\) by repeated application of either \((1)\) or \((-1)\).
\end{definition}
\begin{terminology}
	Let \(M,N\in\gMod_\Bbbk\), and let \(i\in\Z\). A \emph{morphism of degree \(i\)} from \(M\) to \(N\) is a strict morphism
	\(M \to N(i)\).
\end{terminology}
\begin{remark}
	This definition makes sense: a morphism of degree \(i\) should send a degree \(n\) homogeneous element of \(M\) to a degree \(n+i\) homogeneous element
	of \(N\), and indeed, the above definition yields
	\[ M^n \to N^{n+i}. \]
	Given a morphism \(M\to N(i)\) and a morphism \(N\to L(j)\), one can ``compose'' these by forming
	\[ M \to N(i) \to L(i+j), \]
	thus obtaining a morphism from \(M\) to \(L\) of degree \(i+j\).
\end{remark}
\begin{remark}
	Note that since \((i)\) is an automorphism for all \(i\in\Z\), we have that
	\[ \gMod_\Bbbk(M,N) \cong \gMod_{\Bbbk}(M(i),N(i)). \]
	In particular, we get no value from considering maps of the form \(M(i)\to N(j)\), as we can always assume \(i=0\).
\end{remark}

\subsection{Tensor products and Hom in the graded world}
We will now turn to discussing a little bit of actual algebra, in contrast to the above entirely pure definitions. We will define the \emph{tensor product}
of two graded modules in such a way that it is a graded module itself, and similarly produce a graded version of \(\Hom\). For the purposes of the former,
it is useful to introduce \emph{bigraded} modules. These are just modules graded over \(\Z\times\Z\), defined just as one does for \(\Z\).
\begin{definition}
	Let \(\calC\) be a category. The category of \emph{bigraded objects} in \(\calC\) is the functor category
	\[ \Fun(\Z\times\Z,\calC) \cong \prod_{\Z\times\Z}\calC. \]
	In particular, we obtain the category
	\[ \bgMod_{\Bbbk} \coloneq \prod_{\Z\times\Z}\Mod_{\Bbbk}. \]
\end{definition}
The reason we introduce this is because given two graded modules \(M,N\), the most natural way to form a tensor product is to form the
\emph{bigraded} module
\[ (M^i\otimes_\Bbbk N^j), \quad {(i,j)\in\Z\times\Z}. \]
A problem with this, of course, is that we are interested in the world of \emph{graded} modules, not \emph{bigraded} modules. The solution is:
\begin{definition}
	Let \(\calC\) be a category. We define the functor
	\[ \Tot\!:\Fun(\Z\times\Z,\calC)\to\Fun(\Z,\calC),\quad (x^{i,j})_{(i,j)\in\Z\times\Z} \mapsto \left(\coprod_{i+j=n}x^{i,j}\right)_{n\in\Z}.  \]
\end{definition}
The way to visualize this functor is by picturing the bigraded object as lying in a grid, then taking the coproduct of the diagonal lines:
\[
	\begin{tikzcd}
		x^{-1,2}\ar[dr,no head] & x^{0,2}\ar[dr,no head] & x^{1,2}\ar[dr,no head] & x^{2,2} \\
		x^{-1,1}\ar[dr,no head] & x^{0,1}\ar[dr,no head] & x^{1,1}\ar[dr,no head] & x^{2,1} \\
		x^{-1,0}\ar[dr,no head] & x^{0,0}\ar[dr,no head] & x^{1,0}\ar[dr,no head] & x^{2,0} \\
		x^{-1,-1} & x^{0,-1} & x^{1,-1} & x^{2,-1}
	\end{tikzcd}
\]
\begin{proposition}\label{prop:graded-Tot-adjunction}
	Let \(\calC\) be a category admitting countable coproducts. Consider the functor
	\[ G\!: \Fun(\Z,\calC) \to \Fun(\Z\times\Z,\calC),\quad (x^i)_{i\in\Z} \mapsto (x^{i+j})_{(i,j)\in\Z\times\Z}. \]
	Then \(\Tot\ladj G\).
\end{proposition}
\begin{proof}
For \(x\in\Fun(\Z\times\Z,\calC)\) and \(y\in\Fun(\Z,\calC)\), we have
\begin{align*}
	\Hom(\Tot(x),y) &= \prod_{n\in\Z}\Hom(\Tot(x)^n,y^n) \\
	&= \prod_{n\in\Z}\Hom(\coprod_{i+j=n}x^{i,j},y^n) \\
	&\cong \prod_{n\in\Z}\prod_{i+j=n}\Hom(x^{i,j},y^n) \\
	&= \prod_{n\in\Z}\prod_{i+j=n}\Hom(x^{i,j},G(y)^{i,j}) \\
	&= \prod_{(i,j)\in\Z\times\Z}\Hom(x^{i,j},G(y)^{i,j}) = \Hom(x,G(y))
\end{align*}
as desired.
\end{proof}

\begin{definition}
	Let \(M,N\in\gMod_\Bbbk\). Temporarily, let us denote the bigraded \(\Bbbk\)-module \((M^i\otimes_\Bbbk N^j)_{(i,j)\in\Z\times\Z}\) by \(M\otimes_\Bbbk^b N\).
	We define the graded modules
	\begin{align*}
		M\otimes_\Bbbk N &\coloneq \Tot(M\otimes_\Bbbk^b N), \\
		\ugMod_\Bbbk(M,N) &\coloneq (\gMod_\Bbbk(M,N(i)))_{i\in \Z}.
	\end{align*}
	We will also write \(\iHom(M,N) \coloneq \ugMod_\Bbbk(M,N)\) if the context makes it clear what this means. These clearly organize into functors.
\end{definition}

\begin{proposition}\label{prop:graded-tensor-hom-adjunction}
	Let \(M,N,L\in\gMod_\Bbbk\). There is a natural isomorphism
	\[ \gMod_\Bbbk(M\otimes_\Bbbk N, L) \cong \gMod_\Bbbk(M,\ugMod_\Bbbk(N,L)). \]
	In particular, we have an adjunction \(-\otimes_\Bbbk N \ladj \ugMod_\Bbbk(N,-)\).
\end{proposition}
\begin{proof}
Recall the adjunction \(\Tot\ladj G\) from Proposition \ref{prop:graded-Tot-adjunction}. Then
\begin{align*}
	\gMod_\Bbbk(M\otimes_\Bbbk N, L) &= \gMod_{\Bbbk}(\Tot(M\otimes_\Bbbk^b N),L) \\
	&\cong \bgMod_{\Bbbk}(M\otimes_\Bbbk^b N,G(L)) \\
	&= \prod_{(i,j)\in\Z\times\Z}\Mod_{\Bbbk}(M^i\otimes_\Bbbk N^j,L^{i+j}) \\
	&\cong \prod_{(i,j)\in\Z\times\Z}\Mod_{\Bbbk}(M^i,\Mod_\Bbbk(N^j,L^{i+j})) \\
	&= \prod_{i\in\Z}\prod_{j\in\Z}\Mod_{\Bbbk}(M^i,\Mod_\Bbbk(N^j,L^{i+j})) \\
	&\cong \prod_{i\in\Z}\Mod_{\Bbbk}(M^i, \prod_{j\in\Z}\Mod_\Bbbk(N^j,L^{i+j})) \\
	&= \prod_{i\in\Z}\Mod_{\Bbbk}(M^i, \gMod_\Bbbk(N,L(i))) = \gMod_\Bbbk(M,\ugMod_\Bbbk(N,L))
\end{align*}
as desired.
\end{proof}
\begin{remark}
	Note that the foregoing proof and definition goes through in the generality of a closed symmetric monoidal Abelian category. That is,
	an Abelian category \(\calA\) admitting a symmetric monoidal structure \(\otimes\) such that \(-\otimes x\) has a right adjoint for all \(x\in\calA\).
\end{remark}

\begin{remark}
	It is perhaps useful, for purposes of intuition, to observe that the elements homogeneous of degree \(i\) in \(\ugMod_\Bbbk(M,N)\) are exactly the degree \(i\) maps from \(M\) to \(N\).
\end{remark}

\begin{remark}
	By Proposition \ref{prop:graded-Tot-adjunction}, to describe a map \(f\!:M\otimes_\Bbbk N \to L\), it suffices to say how \(f\) acts on elementary tensors
	\[ x\otimes y,\quad x\in M^i,\, y\in N^j, \]
	where one must ensure that \(f(x\otimes y) \in L^{i+j}\).
\end{remark}

\begin{remark}
	The isomorphism \(\gMod_\Bbbk(M\otimes_\Bbbk N,L) \cong \gMod_\Bbbk(M,\ugMod_\Bbbk(N,L))\) is given explicitly by \emph{currying.} That is,
	given a map \(f\!:M\otimes_\Bbbk N \to L\), the corresponding map \(M \to \ugMod_\Bbbk(N,L)\) is defined by the formula
	\[ x \mapsto f(x\otimes-). \]
	Note that if \(x\in M^i\) and \(y\in N^j\), then \(x\otimes y \in (M\otimes_\Bbbk N)^{i+j}\), so \(f(x\otimes y) \in L^{i+j}\), so the morphism
	\(f(x\otimes-)\) does indeed have the expected type \(N \to L(i)\).
\end{remark}

The tensor product \(-\otimes_\Bbbk-\!:\gMod_\Bbbk\times\gMod_\Bbbk\to\gMod_\Bbbk\) endows \(\gMod_\Bbbk\) with the structure
of a (closed) symmetric monoidal category. In particular, it is clear that one has natural isomorphisms \(M\otimes_\Bbbk \Bbbk \cong M\),
and that one has associativity isomorphisms
\[ (M\otimes_\Bbbk N)\otimes_\Bbbk L \cong M\otimes_\Bbbk (N\otimes_\Bbbk L). \]
It is maybe a \emph{little} less clear that these satisfy the various coherences required of a monoidal category, but the idea is that
they are inherited from the ones on \(\Mod_\Bbbk\). This justifies the assertion that we have a \emph{monoidal} structure.

There is a slight complication in how we make the monoidal structure \emph{symmetric.} In particular, the natural isomorphisms
\[ \tau_{M,N}\!: M\otimes_\Bbbk N \iso N\otimes_\Bbbk M \]
are chosen to act by the \emph{Koszul sign rule,}
\[ \tau(x\otimes y) = (-1)^{ij} y\otimes x, \quad x\in M^i,\, y\in N^j. \]
This sign convention is the fundamental reason that graded commutativity is different from ordinary commutativity. The reason one is
interested in this form of commutativity is that one would like objects like the exterior algebra \(\bigwedge^\bullet M\) of a module
\(M\) to be examples of \emph{graded (commutative) algebras,} since these arise often in e.g.\ algebraic geometry and topology. These
exterior algebras satisfy \(x\wedge y = (-1)^{ij}y\wedge x\).

\begin{exercise}
	Show that there is a natural isomorphism
	\[ \ugMod_\Bbbk(M\otimes N,L) \cong \ugMod_\Bbbk(M,\ugMod_\Bbbk(N,L)). \]
	Hint: this is purely formal, and holds in any closed monoidal category.
\end{exercise}
\begin{exercise}
	Check that \(\gMod_\Bbbk\) is an Abelian category.
\end{exercise}

\subsection{Tensor products of morphisms}
We finish our present discussion of the tensor product by explaining how the Koszul sign rule affects \emph{evaluation} of tensor products of morphisms.
The adjunction from Proposition \ref{prop:graded-tensor-hom-adjunction} induces a counit, and it is the \emph{evaluation map}
\[ \ev\!:\ugMod_{\Bbbk}(M,N)\otimes_\Bbbk M \to N,\quad \ev(f\otimes x) = f(x). \]
Note that the formula on the right is slightly more subtle than it may first appear: the evaluation map is a map of \emph{graded modules,} and so
applies to morphisms of non-zero degree as well. More precisely, if \(f\in\ugMod_\Bbbk(M,N)^i\) and \(x\in M^j\), then \(\ev(f\otimes x) = f(x)\in N^{i+j}\),
as expected, since \(f\) is the same as a degree zero map \(M \to N(i)\).

When we have two strict morphisms \(f\!:M \to L\), \(g\!:N\to K\), the morphism \(f\otimes g\) we obtain is simply given by
\[ (f\otimes g)(x\otimes y) = f(x)\otimes g(y). \]
However, what we would like is a way to take tensor products of morphisms of non-zero degree---this does not come for free. Sure, we can
regard a degree \(i\) morphism \(M \to L\) as a strict morphism \(M \to L(i)\) and take the tensor product that way, but this produces something
of the wrong type (at least visually). What we want is a morphism
\[ \ugMod_\Bbbk(M,L) \otimes_\Bbbk \ugMod_\Bbbk(N,K) \to \ugMod_\Bbbk(M\otimes_\Bbbk N, L\otimes_\Bbbk K) \]
taking the ``formal'' tensor product of two arbitrary degree maps to its ``realization'' as an actual morphism. Furthermore, we want this
to be suitably compatible with the evaluation; to be more precise, using the associativity and symmetry isomorphisms together with the evaluation, we can produce
the composite
\[
	\begin{tikzcd}
		(\ugMod_\Bbbk(M,L) \otimes_\Bbbk \ugMod_\Bbbk(N,K))\otimes_\Bbbk (M\otimes_\Bbbk N) \ar[d,"\cong"] \\
		\ugMod_\Bbbk(M,L) \otimes_\Bbbk (\ugMod_\Bbbk(N,K) \otimes_\Bbbk M) \otimes_\Bbbk N \ar[d,"1\otimes\tau\otimes 1"',"\cong"] \\
		\ugMod_\Bbbk(M,L) \otimes_\Bbbk ( M\otimes_\Bbbk \ugMod_\Bbbk(N,K)) \otimes_\Bbbk N \ar[d,"\cong"] \\
		(\ugMod_\Bbbk(M,L) \otimes_\Bbbk M) \otimes_\Bbbk (\ugMod_\Bbbk(N,K) \otimes_\Bbbk N)\ar[d,"\ev\otimes\ev"'] \\
		L\otimes_\Bbbk K
	\end{tikzcd}
\]
and we would like for this to be the map which transposes, via currying, to the ``realization'' morphism
\[ \ugMod_\Bbbk(M,L) \otimes_\Bbbk \ugMod_\Bbbk(N,K) \to \ugMod_\Bbbk(M\otimes_\Bbbk N, L\otimes_\Bbbk K). \]
In the end, this means that we are forced to make the following definition:
\begin{definition}
	Let \(f\in\ugMod_\Bbbk(M,L)\) and let \(g\in\ugMod_\Bbbk(N,K)^i\). We define their tensor product \(f\otimes g\) to be the
	degree \(\deg(f) + i\) morphism \(M\otimes_\Bbbk N \to L\otimes_\Bbbk K\) given by
	\[ (f\otimes g)(x\otimes y) = (-1)^{ij}f(x)\otimes g(y),\quad x\in M^j. \]
\end{definition}
\begin{remark}
	The sign depending on the degrees of \(g\) and \(x\) appearing in the definition comes directly from the Koszul sign rule, since we apply the isomorphism \(\tau\)
	in the above composite.
\end{remark}

\subsection{Graded algebras}
In the context of any symmetric monoidal category, one can define \emph{monoid} objects. Applying that to our situation, we end up with
\begin{definition}
	A \emph{graded \(\Bbbk\)-algebra} is a pair \((A,\mu)\) of a graded \(\Bbbk\)-module \(A\) and a multiplication map
	\[ \mu\!: A\otimes_\Bbbk A\to A \]
	which is associative and unital. That is, there is an element \(1 = 1_A \in A^0\) such that for all \(x\in A^i\), we have \(\mu(1\otimes x) = x = \mu(x\otimes 1)\),
	and one should have \(\mu(\mu(x\otimes y)\otimes z) = \mu(x\otimes\mu(y\otimes z))\). We write \(xy \coloneq \mu(x\otimes y)\).
\end{definition}
\begin{remark}
	As before, it is helpful to note that a map \(A\otimes_\Bbbk A \to A\) consists of the data of maps
	\[ A^i\otimes_\Bbbk A^j \to A^{i+j},\quad \forall i,j\in\Z. \]
	The conditions we require on \(\mu\) are just such that
	\[ 1\cdot x = x = x\cdot 1,\quad (xy)z = x(yz). \]
	That is, the expected axioms.
\end{remark}
\begin{definition}
	We say that a graded algebra \(A\) is \emph{(graded) commutative} if the diagram
	\[
	\begin{tikzcd}
		A\otimes A \ar[r,"\mu"]\ar[d,"\tau","\cong"'] & A \ar[d,equal] \\
		A\otimes A \ar[r,"\mu"] & A
	\end{tikzcd}
	\]
	commutes. Concretely, this means that
	\[ xy = (-1)^{ij}yx,\quad x\in A^i,\, y\in A^j. \]
\end{definition}
\begin{remark}
	Note that the Koszul sign rule means that commutativity, in characteristic not equal to \(2\), necessarily means that
	\[ y^2 = 0 \]
	whenever \(y\) is homogeneous of odd degree. Indeed,
	\[ y^2 = (-1)^{\deg(y)\deg(y)}y^2 = -y^2 \implies 2y^2 = 0. \]
	For this reason, one often \emph{requires} that this holds as part of the definition in the even characteristic case. We will not make
	this distinction here, because we will often implicitly assume we are not working in a context where there will be problems.
\end{remark}
\begin{example}
	We now make precise the example from earlier. Let \(M\) be a \(\Bbbk\)-module. We define \(M\wedge_\Bbbk M\) to be the
	\(\Bbbk\)-module quotient
	\[ M\wedge_\Bbbk M \coloneq (M\otimes_\Bbbk M)/(x\otimes y + y\otimes x)_{x,y\in M}. \]
	We denote the image of \(x\otimes y\) in \(M\wedge_\Bbbk M\) by \(x\wedge y\). The definition is such that \(x\wedge y = -y\wedge x\).

	Now, for all \(k\geq 0\), we define
	\[ \bigwedge^kM \coloneq \underbrace{M\wedge_\Bbbk \cdots \wedge_\Bbbk M}_{k\text{ copies}}, \]
	and set \(\bigwedge^0M \coloneq \Bbbk\). We then have a graded \(\Bbbk\)-module defined by
	\[ \left(\bigwedge M\right)^i = \begin{cases}
		\bigwedge^i M & \text{if }i\geq 0, \\
		0 & \text{if }i \leq 0.
	\end{cases} \]
	This graded module can be endowed with the structure of a graded \(\Bbbk\)-algebra, which is furthermore commutative. In particular, given
	elementary generators \(x = x_1\wedge \cdots \wedge x_i\) and \(y = y_1\wedge \cdots \wedge y_j\), one defines
	\[ x\wedge y = x_1\wedge \cdots \wedge x_i \wedge y_1 \wedge \cdots \wedge y_j \]
	and one easily sees that this extends to maps
	\[ \wedge\!: \left(\bigwedge M\right)^i \otimes_\Bbbk \left(\bigwedge M\right)^j \to \left(\bigwedge M\right)^{i+j} \]
	and thus to yields a map
	\[ \bigwedge M \otimes_\Bbbk \bigwedge M \to \bigwedge M. \]
	It is left as an exercise to check that this is associative and unital. That it is graded commutative comes from the definition of the wedge.
\end{example}

\begin{example}
	Standard examples of commutative \(\Bbbk\)-algebras are polynomial rings over \(\Bbbk\), i.e.\ the \emph{free} commutative algebras over \(\Bbbk\).
	In the graded world, there are similar equivalents, but with the grading comes the freedom to choose what \emph{cohomological degree}---i.e.\ what degree in the grading---the generators
	have.

	Let us first consider the case of a single generator \(x\), yielding a commutative graded \(\Bbbk\)-algebra \(\Bbbk[x]\). If we decide that \(x\) is to be of degree zero,
	then this is just the ordinary polynomial ring in one variable over \(\Bbbk\). Otherwise, the description of \(\Bbbk[x]\) splits into two cases depending on whether the
	degree \(\ell\) of \(x\) is even or odd, and positive or negative. Since the difference between negative and positive isn't very interesting, let us consider
	the case when \(\ell < 0\).

	If \(\ell\) is even, then \(\Bbbk[x]\) as a graded \(\Bbbk\)-module looks like this:
	\[
	\begin{tikzcd}
		\cdots & \Bbbk \cdot x^{k} & 0 & \cdots & \Bbbk \cdot x & 0 & \cdots & \Bbbk & 0 & \cdots
	\end{tikzcd}
	\]
	where a copy of \(\Bbbk\) is placed in positions \(k\ell\) for all \(k \leq 0\), and everything else is zero. The multiplication is such that \(x^{i}x^j = x^{i+j}\).
	In particular, \(\Bbbk[x]\) has an infinite number of copies of \(\Bbbk\) in it.

	If \(\ell\) is odd, then \(\Bbbk[x]\) is considerably simpler. It has the following form:
	\[
	\begin{tikzcd}
		\cdots & 0 & \Bbbk\cdot x & 0 & \cdots & 0 & \Bbbk & 0 & \cdots
	\end{tikzcd}
	\]
	where there is a copy of \(\Bbbk\) placed in degree 0 and degree \(\ell\). This is because we necessarily have \(x^2 = 0\), so all other copies of \(\Bbbk\) die.
\end{example}
\begin{exercise}
	Investigate the commutative graded algebra \(\Bbbk[x,y]\), where we set \(\deg(x) = 0\) and \(\deg(y)=-1\). More precisely, consider the following approaches:
	\begin{enumerate}[label=(\arabic*)]
		\item Describe \(\Bbbk[x,y]\) by hand first, as is done in the above example.
		\item Show that if \(M\) is a graded module concentrated in degree zero (i.e.\ just an ordinary module) and \(N\) is a graded
		module, then \(M\otimes_\Bbbk N\) is the graded module \((M^0\otimes_\Bbbk N^i)_{i\in\Z}\).
		\item Reason that \(\Bbbk[x,y] \cong \Bbbk[x]\otimes_\Bbbk\Bbbk[y]\) in order to describe \(\Bbbk[x,y]\).
		\item Reason that \(\Bbbk[x,y] \cong \Bbbk[x][y]\) in order to describe \(\Bbbk[x,y]\).
	\end{enumerate}
\end{exercise}

\subsection{Appendix: Modules over graded algebras (TBD)}
See \cite[Lemma 3.1.30]{doi:10.1017/9781108292825}.

Recall that we have fixed a commutative ring \(\Bbbk\). Now, let us fix a graded \(\Bbbk\)-algebra \(A\). What does it mean to have a graded \(A\)-module?
\begin{definition}
	A \emph{graded left \(A\)-module} is a graded \(\Bbbk\)-module \(M\) together with a left action of \(A\)
	\[ A\otimes_\Bbbk M \to M,\quad a\otimes m \mapsto am, \]
	which is unital and associative. Dually, a \emph{graded right \(A\)-module} is a graded \(\Bbbk\)-module \(M\) together with a \emph{right} action of \(A\)
	\[ M\otimes_\Bbbk A \to M, \quad m\otimes a \mapsto ma. \]
	Let \(M,N\) be graded left \(A\)-modules. A morphism \(f\!:M\to N\) of graded \(\Bbbk\)-modules is \emph{\(A\)-linear} if
	\[
	\begin{tikzcd}
		A\otimes_\Bbbk M \ar[d]\ar[r,"1\otimes f"] & A\otimes_\Bbbk N\ar[d] \\
		M \ar[r,"f"] & N
	\end{tikzcd}
	\]
	commutes. The definition for graded right \(A\)-modules is similar. We obtain the categories \(\prescript{}{A}{\gMod}\) and \(\gMod_A\).
\end{definition}
\begin{remark}
	Let \(M\) be a graded left \(A\)-module. Note that, by definition, we have
	\[ a\in A^i,\, m\in M^j \implies am\in M^{i+j} \]
	since the action \(A\otimes M \to M\) restricts to \(A^i \otimes_\Bbbk M^j \to M^{i+j}\). A similar statement is true for right \(A\)-modules.
\end{remark}
\begin{remark}
	Note that the definition of \(A\)-linearity makes sense for a morphism of any degree. However, there is a perhaps odd technical detail here:
	if \(f\) is a graded \(\Bbbk\)-module morphism of degree \(i\) between graded \(A\)-modules, then the concrete equations required for \(f\) to be \(A\)-linear
	depends on whether we are working with left or right modules. The reason is the Koszul sign rule. If we are in the case of right graded \(A\)-modules,
	then everything is easy since \((f\otimes 1)(x\otimes a) = f(x)\otimes a\), so we require that \(f(xa) = f(x)a\). On the other hand, for
	\emph{left} graded \(A\)-modules, we are working with \(1\otimes f\), and therefore we require that
	\[ f(ax) = (-1)^{ij}af(x),\quad a\in A^j. \]
\end{remark}
\begin{remark}
	A graded \(\Bbbk\)-algebra \(A\) induces a related graded \(\Bbbk\)-algebra \(A^\op\), whose multiplication is ``opposite'' to the one in \(A\). In particular, one
	applies the symmetry isomorphism to define a new algebra structure by
	\[ *\!:A\otimes_\Bbbk A \overset\tau\to A\otimes_\Bbbk A \overset{\text{mult.}}\to A,\quad \text{i.e.}\quad x*y = (-1)^{ij}yx,\quad x\in A^i,\, y\in A^j. \]
	There is an equivalence of categories \(\prescript{}{A}{\gMod} \simeq \gMod_{A^\op}\); a graded left \(A\)-module \(M\)
	can be endowed with a right \(A^\op\) action induced by the morphism
	\[ M \otimes_\Bbbk A \overset\tau\to A\otimes_\Bbbk M \overset{\text{act.}}\to M. \]
	Notably, we need only ever consider graded right \(A\)-modules. Observe that if \(A\) is graded commutative then \(A=A^\op\).
\end{remark}
\begin{remark}
	For a graded left \(A\)-module \(M\), the action \(A\otimes_\Bbbk M \to M\) transposes by adjunction to a morphism
	\[ A \to \ugMod_\Bbbk(M,M) \eqcolon \underline{\End}_\Bbbk(M). \]
	This yields an equivalent way of thinking about graded left \(A\)-modules. Similarly, a graded right \(A\)-module structure can be encoded
	by a map \(A^\op \to \underline{\End}_\Bbbk(M)\).
\end{remark}

What we would like to do is to extend the tensor-Hom adjunction for graded \(\Bbbk\)-modules to the setting of graded \(A\)-modules. For this,
we need to adapt both the inner Hom \(\ugMod_\Bbbk(-,-)\) and the tensor product \(-\otimes_\Bbbk-\) to their respective \(A\)-linear variants.
The easiest is the former, as we already know what it means for a morphism to be \(A\)-linear:

\begin{definition}
	Let \(A\) be a graded \(\Bbbk\)-algebra, and let \(M,N\in\gMod_A\). We let
	\[ \ugMod_A(M,N)^i \subseteq \ugMod_\Bbbk(M,N)^i \]
	consist of those morphisms of degree \(i\) which are \(A\)-linear. This defines a graded \(A\)-module \(\ugMod_A(M,N)\), which we may also
	sometimes denote \(\iHom_A(M,N)\).
\end{definition}

The tensor product is a little bit trickier, but in the end, we would basically just like to take the tensor product \(-\otimes_\Bbbk-\) we know
about and impose some further equations on it. We may observe that if \(M\) is a right graded \(A\)-module and \(N\) is a left graded \(A\)-module, we have two maps
\[ \ell,r\!: M\otimes_\Bbbk A\otimes_\Bbbk N \to M\otimes_\Bbbk N \]
namely
\[ \ell = (\text{act.})_M\otimes 1_N,\quad \text{and}\quad r=1_M\otimes(\text{act.})_N. \]
Now, \(\gMod_\Bbbk\) is an Abelian category, as all functor categories with codomain an Abelian category are, so we can take the difference of these two maps.
Concretely,
\[ (\ell-r)(x\otimes a \otimes y) = xa\otimes y - x\otimes ay. \]
With this in mind, it makes sense to make the following definition.
\begin{definition}
	Let \(M\) be a right graded \(A\)-module and let \(N\) be a left graded \(A\)-module. We define their tensor product \(M\otimes_A N\) over \(A\) to be
	\[ M\otimes_A N \coloneq \coker(M\otimes_\Bbbk A\otimes_\Bbbk N \xrightarrow{\ell - r} M\otimes_\Bbbk N). \]
	We will abusively still write \(x\otimes y\) for the equivalence class in the cokernel. This becomes a \emph{graded \(A\)-bimodule} with the action
	\[ a(x\otimes y) = x\otimes ay = xa\otimes y = (x\otimes y)a. \]
\end{definition}
\begin{remark}
	It might help to make the grading explicit. To this end, let us investigate the tensor products of the form
	\[ M \otimes_\Bbbk A \otimes_\Bbbk N. \]
	Arbitrarily, let us pick the bracketing \((M\otimes_\Bbbk A)\otimes_\Bbbk N\), and compute
	\begin{align*}
		((M\otimes_\Bbbk A)\otimes_\Bbbk N)^n &= \coprod_{i+j = n} (M\otimes_\Bbbk A)^i\otimes_\Bbbk N^j \\
		&= \coprod_{i+j = n} \left(\coprod_{p+q=i} M^p\otimes_\Bbbk A^q \right) \otimes_\Bbbk N^j \\
		&\cong \coprod_{i+j = n} \left(\coprod_{p+q=i} M^p\otimes_\Bbbk A^q \otimes_\Bbbk N^j \right) = \coprod_{i+t+j=n} M^i\otimes_\Bbbk A^t \otimes_\Bbbk N^j.
	\end{align*}
	so that the \(n\)th graded piece of \(M\otimes_\Bbbk A\otimes_\Bbbk N\) is obtained by taking the coproduct of objects
	\[ M^i\otimes_\Bbbk A^t\otimes_\Bbbk N^j,\quad i+t+j = n. \]
	In particular, we see that
	\[ (M\otimes_A N)^n = \coker\left( \coprod_{i+t+j=n} M^i\otimes_\Bbbk A^t \otimes_\Bbbk N^j \to \coprod_{p+q=n} M^p\otimes_\Bbbk N^q \right) \]
	where the map has exactly the same formula as before.
\end{remark}

For simplicity, we will now restrict to the case where \(A\) is graded commutative. This is to avoid spending time talking about bimodules.

\begin{proposition}
	We have
	\[ \gMod_A(M\otimes_A N,L) \cong \gMod_A(M,\ugMod_A(N,L)). \]
\end{proposition}